---
title: "Population Data"
author: "Liz"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
library(tidyverse) 
library(tidycensus)
library(sf)
library(tigris) 
options(tigris_use_cache = TRUE)
options(scipen = 999)

dir_1 <-"/Users/eliza/Documents/Data/Capstone"
dir_2 <- "/Users/eliza/Documents/Data/Capstone/ACS Data"
```

**ACS Data: Median HH Income - used as census tract proxy**
```{r}
# # Get the data 
# census_tracts <- get_acs(geography = "tract",
#                          state = "NY",
#                          county = c("New York", "Richmond",
#                                     "Kings", "Queens", "Bronx"),
#                         year = 2012,
#                         survey = "acs5",
#                         variables = c(median_hh_income = "B19013_001"),
#                         geometry = TRUE) %>% 
#                 select(-moe) 


# # sf::st_write(census_tracts, str_glue("{dir_2}/census_tracts.shp")) 
# census_tracts %>% head(5) 

census_tracts <- st_read(paste0(dir_2, "/census_tracts_all/median_hh_income.shp")) %>% 
  st_transform(4326) 
```

**Subway Stations (NYC Open Data) includes Treatment and Control Stations**
These stations will be mapped in Arcgis with above data on HH median income. (Arcgis Method)
```{r}
# Read in Subway Stations info from NYC Open Data 
stations <- st_read(paste0(dir_1, "/Subway_Stations/Stations/"), 
                      layer = "geo_export_6c8e1b91-ae30-4310-8364-86bc2ac4e360") %>%  
  st_transform(4326) 


# Treatment Stations gathered from google sheets, MTA dashboard 
names_treat <- c("South Ferry",
           "Rockaway Park - Beach 116 St",
           "Astoria Blvd", 
           "Coney Island - Stillwell Av", 
           "1st Ave", 
           "Whitehall St", 
           "Rector St", 
           "Bowling Green", 
           "Broad St", 
           "Canal St - Holland Tunnel", 
           "Canal St", 
           "Franklin St",
           "145th St", 
           "55th St", 
           "21st St - Queensbridge", 
           "Roosevelt Island - Main St", 
           "Houston St", 
           "138th St - Grand Concourse", 
           "110th St", 
           "103rd St")

treatment_stations <- stations %>% 
  filter(name %in% names_treat)

# st_write(treatment_stations, str_glue("{dir_1}/Treatment_Stations.shp"))

# Comparison Stations in Sandy Inundation Zones from arcgis
names_control <- c("Beach 105th St", 
               "Beach 90th St", 
               "Canal St", 
               "Broad Channel", 
               "Ocean Pkwy", 
               "Beach 44th St", 
               "Beach 60th St", 
               "Beach 98th St",
               "Rockaway Park - Beach 116 St", 
               "Beach 36th St", 
               "W 8th St - NY Aquarium",
               "Beach 67th St",
               "E 105th St",
               "Howard Beach - JFK Airport", 
               "25th Ave",
               "Brighton Beach",
               "Neptune Ave",
               "Ave X",
               "Bay 50th St",
               "Smith - 9th Sts",
               "Canal St - Holland Tunnel",
               "South Ferry",
               "Whitehall St",
               "Franklin St",
              "Coney Island - Stillwell Av") 
               
control_stations <- stations %>% 
  filter(name %in% names_control) %>% 
  filter(!row_number() %in% c(1, 23, 24, 27)) 
               
               
# st_write(control_stations, str_glue("{dir_1}/Control_Stations.shp"))  
```

**Read in treatment census tracts** 
**Hold on control census tracts** 
```{r}
# census tracts may show up more than once because of overlapping buffers in treatment samples (same for control group)

census_tract_denominator <- read_csv(str_glue("{dir_1}/Arcgis_Output/treatment_census_tracts_corrected/census_tracts_denominator.csv")) %>%
  select(-c(1, 4:8)) %>% 
  distinct(GEOID, .keep_all = TRUE)

trt_census_tract_1 <- read_csv(str_glue("{dir_1}/Arcgis_Output/treatment_census_tracts_corrected/trt_census_tracts_1.csv")) %>% 
  mutate(status = "Treatment") %>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)

trt_census_tract_75 <- read_csv(str_glue("{dir_1}/Arcgis_Output/treatment_census_tracts_corrected/trt_census_tracts_75.csv")) %>% 
  mutate(status = "Treatment") %>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)
  

trt_census_tract_50 <- read_csv(str_glue("{dir_1}/Arcgis_Output/treatment_census_tracts_corrected/trt_census_tracts_50.csv")) %>% 
  mutate(status = "Treatment")%>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)

trt_census_tract_25 <- read_csv(str_glue("{dir_1}/Arcgis_Output/treatment_census_tracts_corrected/trt_census_tracts_25.csv")) %>% 
  mutate(status = "Treatment")%>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)
  

## Comment out Control Group Code for now 

census_tract_denominator_ctrl <- read_csv(str_glue("{dir_1}/Arcgis_Output/control_census_tracts_corrected/census_tracts_denominator.csv")) %>%
  select(-c(1, 4:8)) %>% 
  distinct(GEOID, .keep_all = TRUE)

ctrl_census_tract_1 <- read_csv(str_glue("{dir_1}/Arcgis_Output/control_census_tracts_corrected/ctrl_census_tracts_1.csv")) %>%
  mutate(status = "Control")  %>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Length", "SHAPE_Area"))

ctrl_census_tract_75 <- read_csv(str_glue("{dir_1}/Arcgis_Output/control_census_tracts_corrected/ctrl_census_tracts_75.csv")) %>%
  mutate(status = "Control") %>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)
  
ctrl_census_tract_50 <- read_csv(str_glue("{dir_1}/Arcgis_Output/control_census_tracts_corrected/ctrl_census_tracts_50.csv")) %>%
  mutate(status = "Control") %>% 
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)

ctrl_census_tract_25 <- read_csv(str_glue("{dir_1}/Arcgis_Output/control_census_tracts_corrected/ctrl_census_tracts_25.csv")) %>%
  mutate(status = "Control") %>%
  distinct(GEOID, .keep_all = TRUE) %>% 
  select(-c("OBJECTID", "SHAPE_Leng", "SHAPE_Area")) %>% 
  rename(SUM_CT_Area_Clip = SUM_CT_Are)

```


**Calculate Percentage of Census Tract that falls within the area** -- I use threshold criteria of 50%, 75%, and 100% as a sensitivity check 
```{r}

########################### 50% Threshold - TRT ##############################
mile_percent_threshold_50 <- trt_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 50.0) 


three_quarter_mile_percent_threshold_50 <- trt_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 

half_mile_percent_threshold_50 <- trt_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 
  

quarter_mile_percent_threshold_50 <- trt_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 
  

########################### 50 Threshold - CONTROL ########################## 
mile_percent_ctrl_threshold_50 <- ctrl_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 50.0) 

three_quarter_mile_percent_ctrl_threshold_50 <- ctrl_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 


half_mile_percent_ctrl_threshold_50 <- ctrl_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 
  
quarter_mile_percent_ctrl_threshold_50 <- ctrl_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 50.0) 



########################### 75% Threshold - TRT ##############################
mile_percent_threshold_75 <- trt_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 75.0) 


three_quarter_mile_percent_threshold_75 <- trt_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 

half_mile_percent_threshold_75 <- trt_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 
  

quarter_mile_percent_threshold_75 <- trt_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 
  

########################### 75% Threshold - CONTROL ########################## 
mile_percent_ctrl_threshold_75 <- ctrl_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 75.0) 

three_quarter_mile_percent_ctrl_threshold_75 <- ctrl_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 


half_mile_percent_ctrl_threshold_75 <- ctrl_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 
  
quarter_mile_percent_ctrl_threshold_75 <- ctrl_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 75.0) 




########################### 95% Threshold - TRT ##############################
mile_percent_threshold_95 <- trt_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 95.0) 


three_quarter_mile_percent_threshold_95 <- trt_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 

half_mile_percent_threshold_95 <- trt_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 
  

quarter_mile_percent_threshold_95 <- trt_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 
  

########################### 95% Threshold - CONTROL ########################## 
mile_percent_ctrl_threshold_95 <- ctrl_census_tract_1 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = SUM_CT_Area_Clip/CT_Area *100 ) %>% 
  filter(Percent_Area >= 95.0) 

three_quarter_mile_percent_ctrl_threshold_95 <- ctrl_census_tract_75 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 


half_mile_percent_ctrl_threshold_95 <- ctrl_census_tract_50 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 
  
quarter_mile_percent_ctrl_threshold_95 <- ctrl_census_tract_25 %>% 
  left_join(census_tract_denominator, by = "GEOID") %>% 
  mutate(Percent_Area = (SUM_CT_Area_Clip/CT_Area) *100 ) %>% 
  filter(Percent_Area >= 95.0) 

```


**Construct Samples -- only relevant for including an actual control group ** 
```{r}
# census tracts may show up more than once --> once in trt group, once in control group
# filter out these intersecting census tracts

# # .25 Mile Buffer Zone Sample 
quarter_mile_sample <- trt_census_tract_25 %>%
  rbind(ctrl_census_tract_25) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>% # filter on v
  ungroup()

quarter_mile_sample %>% group_by(status) %>% summarize(count = n())

#.50 Mile Buffer Zone Sample
half_mile_sample <- trt_census_tract_50 %>%
  rbind(ctrl_census_tract_50) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

half_mile_sample %>% group_by(status) %>% summarize(count = n())

# .75 Mile Buffer Zone Sample
three_quarter_mile_sample <- trt_census_tract_75 %>%
  rbind(ctrl_census_tract_75) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

three_quarter_mile_sample %>% group_by(status) %>% summarize(count = n())

# 1 Mile Buffer Zone Sample
one_mile_sample <- trt_census_tract_1 %>%
  rbind(ctrl_census_tract_1) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

one_mile_sample %>% group_by(status) %>%
  summarize(count = n())

################################  SAMPLES at 50 % threshold  #######################
# # .25 Mile Buffer Zone Sample 
quarter_mile_sample_threshold_50 <- quarter_mile_percent_threshold_50 %>%
  rbind(quarter_mile_percent_ctrl_threshold_50) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>% # filter on v
  ungroup()

quarter_mile_sample_threshold_50 %>% group_by(status) %>% summarize(count = n())

#.50 Mile Buffer Zone Sample
half_mile_sample_threshold_50 <- half_mile_percent_threshold_50 %>%
  rbind(half_mile_percent_ctrl_threshold_50) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

half_mile_sample_threshold_50 %>% group_by(status) %>% summarize(count = n())

# .75 Mile Buffer Zone Sample
three_quarter_mile_sample_threshold_50 <- three_quarter_mile_percent_threshold_50 %>%
  rbind(three_quarter_mile_percent_ctrl_threshold_50) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

three_quarter_mile_sample_threshold_50 %>% group_by(status) %>% summarize(count = n())

# 1 Mile Buffer Zone Sample
one_mile_sample_threshold_50 <- mile_percent_threshold_50 %>%
  rbind(mile_percent_ctrl_threshold_50) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

one_mile_sample_threshold_50 %>% group_by(status) %>%
  summarize(count = n())

################################  SAMPLES at 75 % threshold  #######################
# # .25 Mile Buffer Zone Sample 
quarter_mile_sample_threshold_75 <- quarter_mile_percent_threshold_75 %>%
  rbind(quarter_mile_percent_ctrl_threshold_75) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>% # filter on v
  ungroup()

quarter_mile_sample_threshold_75 %>% group_by(status) %>% summarize(count = n())

#.50 Mile Buffer Zone Sample
half_mile_sample_threshold_75 <- half_mile_percent_threshold_75 %>%
  rbind(half_mile_percent_ctrl_threshold_75) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

half_mile_sample_threshold_75 %>% group_by(status) %>% summarize(count = n())

# .75 Mile Buffer Zone Sample
three_quarter_mile_sample_threshold_75 <- three_quarter_mile_percent_threshold_75 %>%
  rbind(three_quarter_mile_percent_ctrl_threshold_75) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

three_quarter_mile_sample_threshold_75 %>% group_by(status) %>% summarize(count = n())

# 1 Mile Buffer Zone Sample
one_mile_sample_threshold_75 <- mile_percent_threshold_75 %>%
  rbind(mile_percent_ctrl_threshold_75) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

one_mile_sample_threshold_75 %>% group_by(status) %>%
  summarize(count = n())



################################  SAMPLES at 95 % threshold  #######################
# # .25 Mile Buffer Zone Sample 
quarter_mile_sample_threshold_95 <- quarter_mile_percent_threshold_95 %>%
  rbind(quarter_mile_percent_ctrl_threshold_95) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>% # filter on v
  ungroup()

quarter_mile_sample_threshold_95 %>% group_by(status) %>% summarize(count = n())

#.50 Mile Buffer Zone Sample
half_mile_sample_threshold_95 <- half_mile_percent_threshold_95 %>%
  rbind(half_mile_percent_ctrl_threshold_95) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

half_mile_sample_threshold_95 %>% group_by(status) %>% summarize(count = n())

# .75 Mile Buffer Zone Sample
three_quarter_mile_sample_threshold_95 <- three_quarter_mile_percent_threshold_95 %>%
  rbind(three_quarter_mile_percent_ctrl_threshold_95) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

three_quarter_mile_sample_threshold_95 %>% group_by(status) %>% summarize(count = n())

# 1 Mile Buffer Zone Sample
one_mile_sample_threshold_95 <- mile_percent_threshold_95 %>%
  rbind(mile_percent_ctrl_threshold_95) %>%
  group_by(GEOID) %>%
  filter(!(n() >1)) %>%
  ungroup()

one_mile_sample_threshold_95 %>% group_by(status) %>%
  summarize(count = n())



```


```{r}
# SAMPLES 
census_tracts_threshold_50 <- quarter_mile_sample_threshold_50 %>% 
  rbind(half_mile_sample_threshold_50) %>% 
  rbind(three_quarter_mile_sample_threshold_50) %>% 
  rbind(one_mile_sample_threshold_50) %>% 
  mutate(area_threshold_flag = "50 percent")


census_tracts_threshold_75 <- quarter_mile_sample_threshold_75 %>% 
  rbind(half_mile_sample_threshold_75) %>% 
  rbind(three_quarter_mile_sample_threshold_75) %>% 
  rbind(one_mile_sample_threshold_75) %>% 
  mutate(area_threshold_flag = "75 percent")
  

census_tracts_threshold_95 <- quarter_mile_sample_threshold_95 %>% 
  rbind(half_mile_sample_threshold_95) %>% 
  rbind(three_quarter_mile_sample_threshold_95) %>% 
  rbind(one_mile_sample_threshold_95) %>% 
  mutate(area_threshold_flag = "95 percent")




census_tracts <- census_tracts_threshold_50 %>% 
  rbind(census_tracts_threshold_75) %>% 
  rbind(census_tracts_threshold_95)

write_csv(census_tracts, str_glue("{dir_2}/census_tracts.csv")) 
```




